\begin{thebibliography}{}

\bibitem{an14}
\newblock An N., Mittag J. and H. Hartenstein, 
\newblock "Designing fail-safe and traffic efficient 802.11p-based rear-end collision avoidance," in 2014 IEEE Vehicular Networking Conference (VNC), 2014, pp. 9–16.

\bibitem{anderson79}
\newblock Anderson B. and Moore J.,
\newblock "Optimal Filtering",
\newblock Prentice-Hall, Englewood Cliffs, New Jersey, 1979.

\bibitem{antsaklis07}
\newblock Antsaklis P. J. and Michel A. N.,
\newblock "A Linear Systems Primer",
\newblock 2007 edition, Birkhauser Boston, 2007.
 

\bibitem{biswas09}
\newblock Biswas S., Tatchikou R. and F. Dion, 
\newblock "Vehicle-to-vehicle wireless communication protocols for enhancing highway traffic safety,"
\newblock IEEE Communications Magazine, vol. 44, no. 1, pp. 74–82, Jan. 2006.

\bibitem{brown06}
\newblock Brown F. T.,
\newblock "Engineering System Dynamics. A Unified Graph-Centered Approach",
\newblock Second Edition, CRC Press, 2006. 

\bibitem{brunson02}
\newblock Brunson S.J. et al., 
\newblock "Alert Algorithm Development Program. NHTSA Rear-End Collision Alert Algorithm. Final Report", 
\newblock Technical Report DOT HS 809 526 U.S. Department of Transportation National Highway Traffic Safety Administration, September 2002.

\bibitem{candy09}
\newblock Candy J. V.,
\newblock "Bayesian Signal Processing: Classical, Modern and Particle Filtering Methods", 
\newblock New Jersey: John Wiley and Sons, Inc., 2009.

\bibitem{edstrand12}
\newblock Edstrand J.,
\newblock "Calculation method for powering a tramway network",
\newblock Master's thesis, Chalmers university of technology, 2012.

\bibitem{franklin18}
\newblock Franklin F. G., Powell J.D. and Emami-Naeini, Abbas.
\newblock "Feedback Control of Dynamic Systems", 
\newblock ISBN 978-1-29-206890-6, NY, Pearson, Eighth edition, 2018.

\bibitem{heirich2011}
\newblock Heirich, O., A. Lehner, P. Robertson, and T. Strang,
\newblock "Measurement and Analysis of Train Motion and Railway Track Characteristics with Inertial Sensors.",
\newblock in Proc. 14th Int. IEEE Conference on Intelligent Transportation Systems, 2011, pp. 1995–2000.

\bibitem{its14}
\newblock Intelligent Transport Systems (ITS), 
\newblock Vehicular communications, 
\newblock Basic set of applications, 
\newblock Part 2: Specification of cooperative awareness basic service, 2014.

\bibitem{ISO_geo}
\newblock ISO 6709:2008.
\newblock "Standard representation of geographic point location by coordinates."

\bibitem{ISO_trans}
\newblock ISO 15623:2013.
\newblock "Intelligent transport systems -- Forward vehicle collision warning systems -- Performance requirements and test procedures."

\bibitem{karnopp12}
\newblock Karnopp D. C., Margolis D. L., Rosenberg R. C.,
\newblock "System Dynamics: Modeling and Simulation of Mechatronic Systems.",
\newblock ISBN: 978-0-470-88908-4, Wiley; Fifth edition, 2012.

\bibitem{mukhtar15}
\newblock Mukhtar, A., L. Xia, and T. B. Tang.,
\newblock "Vehicle Detection Techniques for Collision Avoidance Systems: A Review.",
\newblock IEEE Transactions on Intelligent Transportation Systems 16, no. 5 (October 2015): 2318–38. https://doi.org/10.1109/TITS.2015.2409109.

\bibitem{nilsson98}
\newblock Nilsson J.,
\newblock Real-Time Control Systems with Delays,
\newblock Doctoral thesis, Department of Automatic Control, Lund Institute of Technology, 1998

\bibitem{pichlik18}
\newblock Pichlík P., 
\newblock "Strategy of Railway Traction Vehicles Wheel Slip Control",
\newblock Doctoral Thesis, Czech technical university in Prague, 2018.

\bibitem{reid01}
\newblock Reid I., 
\newblock "Estimation II, Lecture notes",
\newblock Available online:\\ \href{robots.ox.ac.uk:5000/\~ian/Teaching/Estimation/LectureNotes2.pdf}{http://www.robots.ox.ac.uk:5000/\~ian/Teaching/Estimation/LectureNotes2.pdf}, \\ 
\newblock Hilary Term, 2001.

\bibitem{sadr16}
\newblock Sadr S., Khaburi D. A. and Rodríguez J.,
\newblock "Predictive Slip Control for Electrical Trains",
\newblock IEEE Transactions on Industrial Electronics, vol. 63, no. 6, pp. 3446–3457, Jun. 2016.

\bibitem{simon06}
\newblock Simon, D.
\newblock "Optimal State Estimation, Kalman, H$_\infty$ and Nonlinear Approaches."
\newblock John Wiley \& Sons, Inc.m Hoboken, New Jersey, 2006.

\bibitem{titterton04}
\newblock Titterton D.H. and Weston, J.L. 
\newblock "Strapdown Inertial Navigation Technology."
\newblock 2nd Edition, AIAA Education Series, Reston, 2004.
\newblock https://dx.doi.org/10.1049/PBRA017E

\bibitem{willke09}
\newblock Willke T. L., Tientrakool P. and Maxemchuk N. F., 
\newblock "A survey of inter-vehicle communication protocols and their applications",
\newblock IEEE Communications Surveys Tutorials, vol. 11, no. 2, pp. 3–20, Second 2009.

\bibitem{xiang14}
\newblock Xiang X., Qin W. and Xiang B., 
\newblock "Research on a DSRC-Based Rear-End Collision Warning Model",
\newblock "IEEE Transactions on Intelligent Transportation Systems", vol. 15, no. 3, pp. 1054–1065, Jun. 2014.


\end{thebibliography}
