\chapter{Introduction to FOC - work name}
\section{FOC overview}
Field Oriented Control is a closed loop control method for controlling torque of BLDC motors based on Clarke-Park Transformation.
Standard BLDC motor is aquipped with three stator windings, which are supplied with three phase currents.
To achive commutation and generate torque, propper voltage has to be provided to the stator windings by reading the phase currents ($i_a, i_b, i_c$).
Three phase currents reference frames can be decomposed into two orthogonal torque and flux reference frames using Clarke transformation, 
resulting in two currents ($i_\alpha$, $i_\beta$) in a stationary reference frame.

Park transformation is then applied to transform the currents from a stationary reference frame to a rotating reference frame, resulting in two DC currents ($i_d$, $i_q$), 
where $i_d$ is a flux current and $i_q$ is a torque current.

To control $i_d$ and $i_q$ currents, two independent proportional-integral (PI) controllers are used to update applied voltage to the stator windings based on the error between the reference and the measured currents.
In this implementation only the $i_q$ current reference is changed to control the torque, while the $i_d$ current reference is always set to zero to maximize the torque per ampere.

After getting new voltage references from the PI controllers, they need to be transformed back to three phase voltages using inverse Park and inverse Clarke transformations,
which can be then applied to the stator windings.

However applying the voltages directly from the inverse Park-Clarke transformations will achieve only aproximatelly 84 of the maximum voltage difference between the stator windings,
which is not optimal. To maximize the voltage difference between the stator windings, Space Vector Modulation (SVPWM) can be applied to the voltage references. 

Now that we can control the torque of the motor by controlling the $i_q$ current reference, a velocity control loop can be implemented on top of the torque control by 
using a PI controller to update the $i_q$ current reference based on the velocity error.

Aditionally, a position control loop can be implemented on top of the velocity control loop by using a PID controller to update the velocity reference based on the position error.

*add diagram of the control loops*

\section{Clarke transformation}
Clarke transformation is used to transform three phase currents ($i_a$, $i_b$, $i_c$) into two orthogonal currents ($i_\alpha$, $i_\beta$) in a stationary reference frame,
where $i_\alpha$ is torque generating current and $i_\beta$ is flux producing current.

Clarke transformation can be represented as a matrix multiplication of the three phase currents with the following transformation matrix:
\begin{equation}
\begin{bmatrix}
i_\alpha \\
i_\beta
\end{bmatrix}
=
\frac{2}{3}
\begin{bmatrix}
1 & -\frac{1}{2} & -\frac{1}{2} \\
0 & \frac{\sqrt{3}}{2} & -\frac{\sqrt{3}}{2}
\end{bmatrix}
\begin{bmatrix}
i_a \\
i_b \\
i_c
\end{bmatrix}
\end{equation}

The scaling factor $\frac{2}{3}$ ensures amplitude invariance between the three-phase and $i_\alpha$ $i_\beta$ representations.

Inverse Clarke transformation in context of FOC is used to transform two orthogonal voltages ($v_\alpha$, $v_\beta$) in a stationary reference frame back
to three phase voltages ($v_a$, $v_b$, $v_c$) that can be applied to the stator windings.
\begin{equation}
\begin{bmatrix}
v_a \\
v_b \\
v_c
\end{bmatrix}
=
\begin{bmatrix}
1 & 0 \\
-\frac{1}{2} & \frac{\sqrt{3}}{2} \\
-\frac{1}{2} & -\frac{\sqrt{3}}{2}
\end{bmatrix}
\begin{bmatrix}
v_\alpha \\
v_\beta
\end{bmatrix}
\end{equation}


\section{Park transformation}
Park transformation is used to transform two orthogonal currents ($i_\alpha$, $i_\beta$) in a stationary reference frame from the stator point of view
into two DC currents ($i_d$, $i_q$) in a rotating reference frame from a rotor point of view, where $i_d$ is a flux current and $i_q$ is a torque current.
To perfrom Park transformation, the electrical angle $\theta_e$ must to be calculated based on the rotor position, which can be obtained from the encoder reading.

Park transformation can be represented as a matrix multiplication of the two orthogonal currents ($i_\alpha$, $i_\beta$) with the following transformation matrix:

\begin{equation}
\begin{bmatrix}
    i_d \\
    i_q
\end{bmatrix}
=
\begin{bmatrix}
\cos(\theta_e) & \sin(\theta_e) \\
\sin(-\theta_e) & \cos(\theta_e)
\end{bmatrix}
\begin{bmatrix}
    i_\alpha \\
    i_\beta
\end{bmatrix}
\end{equation}

In context of FOC, inverse Park transformation is used to transform two orthogonal DC voltages ($v_d$, $v_q$), which are the outputs of the PI controllers, 
from a rotating reference frame back to two orthogonal voltages ($v_\alpha$, $v_\beta$) in a stationary reference frame.

Inverse park transformation can be represented as a matrix multiplication of the two orthogonal DC voltages ($v_d$, $v_q$) with the following transformation matrix:
\begin{equation}
\begin{bmatrix}
v_\alpha \\
v_\beta
\end{bmatrix}
=
\begin{bmatrix}\cos(\theta_e) & -\sin(\theta_e) \\
\sin(\theta_e) & \cos(\theta_e)
\end{bmatrix}
\begin{bmatrix}v_d \\
v_q
\end{bmatrix}
\end{equation}

\section{PI regulator}
Anti-windup
\section{Space vector modulation}
Space vector modulation is used to maximize the voltage difference between the stator windings.

\section{Velocity and position control loops}
\section{Electrical angle calculation}