\chapter{Future improvements}
While it was demonstrated that the uFOC firmware correctly implemented the FOC algorithm, 
multiple improvements are planned for future firmware iterations.

\section{Execution time of torque control loop}
Execution speed of the torque control loop directly affects the maximum achievable motor speed and overall control bandwidth.
The current implementation operates at a maximum stable frequency of approximately 8.9~kHz, which was sufficient for the tested operating conditions.
However, several optimizations are proposed for future firmware revisions in order to potentially achieve up to twice the current execution speed.

\subsection{Encoder reading limitations}
The main limitation of the current torque control loop execution time is the SPI communication
with the encoder and processing performed inside the ADC callback interrupt.
Measurements showed that the torque loop itself can operate at frequencies up to approximately 20~kHz 
when encoder communication and processing are removed.
However, with encoder updates enabled, the maximum stable loop frequency was measured to be approximately 8.9~kHz.

A possible improvement would be to execute the \texttt{encoder\_update()} function in a 
separate interrupt running at a lower frequency than the torque control loop.
To maintain accurate rotor position estimation for the FOC algorithm, 
electrical angle prediction could be implemented using the previously measured 
electrical angular velocity together with the elapsed time since the last encoder measurement.

Such an approach could significantly reduce the execution time of the main torque control loop
while preserving sufficient angle estimation accuracy for stable motor control.

\subsection{Current sampling optimization}
Current sampling could also be further optimized. In the current implementation, currents of all three motor phases are measured independently.
However, according to Kirchhoff's current law, the sum of all three phase currents in a balanced three-phase system is equal to zero,

\begin{equation}
    i_a + i_b + i_c = 0 
\end{equation}

Therefore, only two phase currents need to be measured directly, while the third current can be calculated from the remaining two currents. \cite{kirchhoff}
This approach could reduce ADC processing overhead, potentially allowing higher control loop frequencies.


\section{Encoder excentricity calibration}
The accuracy of encoder angle measurement depends on how precisely the diametrically magnetized magnet is attached to the motor shaft,
as well as how accurately the center of rotation is aligned with the physical encoder sensing element.
Incorrect alignment may introduce angle measurement errors caused by magnet eccentricity and off-axis misalignment.

The MT6835 encoder includes multiple built-in calibration mechanisms intended to compensate for these errors and improve angular measurement linearity.
According to the datasheet, the encoder supports factory calibration, user auto-calibration and user non-linearity compensation (NLC) calibration~\cite{datasheet_MT6835}. 
The user auto-calibration mode allows the encoder to automatically calculate compensation coefficients for a specific magnetic field configuration and assembly condition,
reducing the integral non-linearity (INL) error to typically less than $\pm0.07^\circ$~\cite{datasheet_MT6835}. 
For even higher precision applications, the encoder also provides a 256-point lookup-table based Non-Linearity Compensation (NLC) calibration mode.

\section{Anticogging}
Cogging torque is an undesirable effect commonly present in low cost BLDC and PMSM motors, caused by the interaction between permanent rotor magnets and stator teeth.
This effect introduces periodic torque ripple, especially noticeable at low rotational speeds, resulting in non smooth motion and degraded position control accuracy.

Two common approaches are used to reduce cogging effects.
The first approach involves mechanical and electromagnetic motor design optimizations, such as skewed stator laminations or optimized magnet geometry.
However, these solutions generally increase manufacturing complexity and cost.

The second approach utilizes software based calibration and compensation methods.
These methods estimate periodic torque disturbances over one mechanical revolution and apply compensation torque during operation.
According to researchers at Modlab UPenn~\cite{anticogging}, actuator cost and torque smoothness are often inversely related, 
meaning that low cost actuators typically exhibit larger torque ripple and cogging effects.
Therefore, software anticogging compensation can significantly improve smoothness of motion even on inexpensive actuators.

For this reason, implementation of anticogging calibration and compensation is proposed as a future firmware improvement.


\section{Multi-motor path synchronization}
The current implementation allows communication with multiple motor drivers over a shared CAN bus line.
However, individual drivers currently operate independently without synchronized trajectory execution.

Path synchronization is a desirable feature for applications involving kinematic systems with multiple degrees of freedom,
where coordinated motion between multiple actuators is required.
For example, in applications such as 3D printers, CNC machines or robotic manipulators,
unsynchronized actuator movement could result in trajectory distortion, where a commanded straight line motion may instead produce a curved path.

Therefore, implementation of multi axis trajectory synchronization is proposed as a future improvement.
Synchronization could be implemented either on the client side by distributing time aligned commands,
or preferably directly between motor drivers using CAN bus communication for deterministic coordinated motion control.

\section{IMU support}
The uFOC hardware design includes an onboard IMU sensor, however IMU support has not yet been implemented in the firmware.
The primary reason is the shared SPI bus configuration with the encoder,
which complicates deterministic high speed communication required by the torque control loop.

Nevertheless, IMU integration is considered an important future improvement.
Availability of inertial measurements could enable advanced features such as orientation estimation,
vibration analysis or resonance compensation techniques such as input shaping used in modern 3D printers~\cite{klipper_resonance}.

\section{Developer feedback}
The firmware has not yet been tested by external developers or users.
Therefore, collecting developer feedback is considered an important future step in the firmware development process.

Practical usage by additional developers could reveal usability issues, undocumented edge cases, missing functionality,
or opportunities for further optimization and API improvements.
Feedback gathered from real world applications may also help prioritize future 
development directions and improve overall firmware robustness and maintainability.
